# Chapter 2: Analysis

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  
In here, I first read the content of the table to variable lrn14 from given URL. The table has information of "Johdatus tilastotieteeseen"- courses survey. Let's look at the overall structure of the table.
```{r setup, echo=FALSE, include=FALSE}
library(dplyr)
lrn14 <- read.csv("data/learning2014.csv")
keep_columns <- c("gender","Age","Attitude", "deep", "stra", "surf", "Points")
lrn14 <- select(lrn14, one_of(keep_columns))
```

```{r structure, echo=FALSE}
str(lrn14)

dim(lrn14)
```
As we can see here, there are seven columns in this table: Gender, Age, Attitude, Deep, Stra, Surf and Points. The values of columns are separated by comma ",".Gender and Age are pretty self- explanatory: age and gender of participant of the survey. Attitude questions asked about participants attitude towards statistics. Deep, Stra and Surf describe the questions given in the survey as being deep, strategic or surface. The last column Points is the points this participant got from the course exam. 

In  the end are the dimensions of the table: 167 entries and 7 variables (=columns). 

```{r overview, echo=FALSE}

p1 <- plot(lrn14)

```


Here is the graphical overview of the table. It is visible in this overview that strategic and surf questions are somehow correlated. The surveys participants are mostly close to age 20 and exam points are fairly evenly distributed.

```{r summary}

summary(lrn14$gender)

summary(lrn14$Age)

summary(lrn14$Attitude)

summary(lrn14$deep)

summary(lrn14$stra)

summary(lrn14$surf)

summary(lrn14$Points)

```

The exact information is easier viewed from the summaries of the columns. In the overview it was hard to determine what was the ratio of genders, but in here we can see that there were more female participants than male. 

And what comes to the results, some questions did get maximum points from the scale (e.g Strategic and Attitude). Most of the mean values are above the mean of the  in total. 
```{r setup2, echo=FALSE, include=FALSE}

library(ggplot2)
library(GGally)
library(dplyr)

```

```{r regression, echo=FALSE}

my_model2 <- lm(Points ~ Attitude, data = lrn14)
summary(my_model2)

```

Above here is the multiple regression models summary. First, let's look at the residuals. The values of residuals look normal distributed, which means that the values read from left to right would make a Gaus curve if drawn. 


The stars on the right side of values represent the significance of the variables. The more stars, the more significant it is. Attitude has clearly a significance whereas surface, strategic and deep had no stars at all. These are not visible, since the description of exercise told to remove them.



The multiple R squared of the model is approx. 0,19. The best value is 1, therefore this model is moderate and means that ~19% of the cause of exam points is due to attitude. Warning: remember that correlation is not always same as causality!

```{r validation, echo=FALSE}
par(mfrow = c(2,2))
plot(my_model2, which = c(1:2,5))
```

The last charts are the diagnostics plots for the model. According to the values seen before, assumption is that the modle is moderate: some values fit the prediction and some do not.The suprising element here is that only chart that fits is Normal QQ.On the other hand, Residuals vs. Fitted and Residuals vs. Leverage resemble the multiple R squared value of ~19%. This conludes that fact that this model is neither good or bad. 

