# Chapter 2: Analysis

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  
In here, I first read the content of the table to variable lrn14 from given URL. The table has information of "Johdatus tilastotieteeseen"- courses survey. Let's look at the overall structure of the table.
```{r setup, echo=FALSE, include=FALSE}
library(dplyr)
lrn14 <- read.csv("data/learning2014.csv")
keep_columns <- c("gender","Age","Attitude", "deep", "stra", "surf", "Points")
lrn14 <- select(lrn14, one_of(keep_columns))
```

```{r structure, echo=FALSE}
str(lrn14)

dim(lrn14)
```
As we can see here, there are seven columns in this table: Gender, Age, Attitude, Deep, Stra, Surf and Points. The values of columns are separated by comma ",".Gender and Age are pretty self- explanatory: age and gender of participant of the survey. Attitude questions asked about participants attitude towards statistics. Deep, Stra and Surf describe the questions given in the survey as being deep, strategic or surface. The last column Points is the points this participant got from the course exam. 

In  the end are the dimensions of the table: 167 entries and 7 variables (=columns). 

```{r overview, echo=FALSE}

p1 <- plot(lrn14)

```


Here is the graphical overview of the table. It is visible in this overview that strategic and surf questions are somehow correlated. The surveys participants are mostly close to age 20 and exam points are fairly evenly distributed.

```{r summary}

summary(lrn14$gender)

summary(lrn14$Age)

summary(lrn14$Attitude)

summary(lrn14$deep)

summary(lrn14$stra)

summary(lrn14$surf)

summary(lrn14$Points)

```

The exact information is easier viewed from the summaries of the columns. In the overview it was hard to determine what was the ratio of genders, but in here we can see that there were more female participants than male. 

And what comes to the results, some questions did get maximum points from the scale (e.g Strategic and Attitude). Most of the mean values are above the mean of the  in total. 
```{r setup2, echo=FALSE, include=FALSE}

library(ggplot2)
library(GGally)
library(dplyr)

```

```{r regression, echo=FALSE}

my_model2 <- lm(Points ~ deep + stra + surf, data = lrn14)
summary(my_model2)

```